### 2020-02-04 ChapterIntroduction/chapterIntroduction.tex
Updated: Re-arrange and rewrite of some sections of Introduction

\chapter{Introduction}

\section{Overview}

%Physics is important for games, it is used for...
Video games rely on real-time physics engines to provide realistic environments. This is noticeable in 3D gaming where players expect fast paced interaction. This requirement has produced significant commercial activity resulting in physics engines for real-time simulation for rigid body dynamics. Although realism is convincing, a degree of mathematical accuracy is usually sacrificed (causing error) in order to attain real-time delivery within constrained resource settings. However, with significant commercial effort, the mathematical error has reduced over the years and commercial game engines are finding their use in other industries (e.g.,  \cite{Xu2017, Lu2017, Shah2018}).

Combining physics engines with game engines (e.g., Unity, Unreal), presents a cost-efficient development platform for developing realistic environments. These gaming environments may be presented to geographically distant players using server-side scalable technologies. This allows scalable gaming platforms, presenting players with large online worlds to explore whilst maintaining the visually highly realistic environments generated by the player's local machine. Research into delivering scalability in such online worlds is focused on balancing real-time and consistency issues for enabling player-player interaction. The evolution of such research can be traced back to the earliest work on game area subdivision \cite{Npsnet} through player-focused game area sub-division (e.g., \cite{Morgan2005InterestManagement, Greenhalgh}) and eventually to the many commercial cloud based solutions as described by \cite{iCloudAccess}.

Scaling and more complex games is a selling point.
The problem of scalable player-player interaction is not a solved problem, as the numerous interaction possibilities across gaming genres and the latency in networks provides many different possible bespoke solutions. Player interaction can be measured in milliseconds, which current networking technology cannot model for all gaming genres in a scalable manner without giving rise to inconsistency. Therefore, the notion of trying to model real-time physics, which requires solvers working in iterations measured in microseconds, would be considered a near impossible task. However, achieving this would allow the physics engine element of the gaming console to be transplanted to the server-side (cloud), freeing up resources for game-play and rendering on a player's local machine.
Traditionally real-time simulations are confined to a single machine, including real-time physics. 
Recent cloud technology has made available, lower cost, low-latency, high-performance, including GPU-enabled machines. This means, real-time simulations can now run in the cloud as well as being economically viable. It is possible to rapidly provision resources on demand, allowing for fast scaling up and down of resources as needed.

Recent developments have demonstrated the importance of software defined networks (SDN) for presenting timely cloud solutions \cite{Wang, Kumar}. The SDN manages the distribution of messages across servers in the most economical way possible (e.g., resource usage). The fact that SDNs can produce timely messaging across tightly coupled server-side deployments (the cloud) suggests the possibility that timely requirements for scalable physics delivery can be fulfilled.

@@

Streamed gaming is an active area of research (e.g., \cite{GARCIAVALLS}), where the game is executed on the server-side, the graphical output is streamed to a remote player and input is sent from the player to the server-side (e.g., \cite{PSNOW, NVidiaGameStream}). The existence of such services suggests the possibility of real-time interaction for all gaming genres across cloud infrastructures.

%Streamed gaming is an active area of research, where the game is run entirely in the cloud, with the graphical output streamed to the player. 
Streamed gaming reduces the computing power requirements on player hardware, even allowing gaming on mobile devices with graphical quality beyond what would be possible if the game were running entirely the device. If the physics simulation is co-located with the graphics streaming servers (with very low-latency), the problem of latency between player's view of the physics simulation and actual state is resolved, allowing for a better shared view of the physics state.

It therefore is the next logical step to start breaking apart the different systems of the game into their own 'containers', allowing for independent scaling and resource provision management. Our problem is concentrated on distributing the physics engine element of game engines across the server-side while maintaining the real-time streamed service to players.
Using the recent developments in cloud technology, we explore the possibility of using this on-demand scaling, to scale up real-time physics simulations.

In the context of current research, we can place ourselves at the juncture of balancing physics calculations across servers in the cloud for game streaming services. This should provide improved economic use of resources, as current game streaming services simply mimic the machine requirements of the player without distributing any of the component parts of the gaming engine. In addition, streamed games may benefit from increased numbers of artefacts (way beyond what is currently possible in a player's machine) as other servers may be utilised to solve the physics problems.

@@

\section{Outline}
The following section provides an outline for this study:
2. Background and Related Work:
\begin{itemize}
	%Context
	\item Real-time physics engines overview. We introduce the reader to the topic of physics simulations and we compare real-time with non-real-time physics simulations. We define terms, workings, important and relevant aspects, e.g. body types, uses in games and simulations, typical scales and performance of real-time simulations and performance budget in real-time applications.
	\item Scalable Non-Real Time Physics. We discuss previous work and applications of scalable non-real time physics and how those techniques are not suitable for use in real-time physics simulations.
	\item Online gaming. We discuss the wider topic of online gaming (client-hosted, P2P, server), challenges of latency, security/cheating and computational limits. Also some details of online gaming servers are discussed, specifically, typical server frame-times and network tick updates. 
	\item Physics in online gaming (and the lack of, due to the challenges faced).
	%Relevant Related Work
	\item Graph-partitioning. When it comes to spatially-partitioning a simulation, the problem is essentially a graph partitioning problem, where objects are nodes and if there’s a possible interaction, an edge exists. Useful for terms, comparisons with and future work on load-balancing will heavily involve this.
	\item Distributed virtual environments. We look at existing ways of distributing game worlds and previous work carried out on these approaches. These fall into two categories, migratory (approach used by AP) and non-migratory. The studies found non-migratory to be the better solution when focusing on player-player interaction and we will discuss why migratory is the better solution for physics body-physics body interaction. Important concepts such as interest management are introduced to the reader.
	\item Distributed Real-time Physics. We will discuss previous work looking specifically at distributing real-time physics and how AP differs from these approaches.
	\item Cloud Computing. We discuss the wider context of cloud computing, how AP is deployed on the cloud and also specifically conditions that affect AP in the cloud, i.e. latency and packet-loss
	%Applications
	\item Scientific and Engineering simulations. Real-time simulations are being used more in scientific research, however these are limited to what can be achieved on a single machine. AP allows real-time physics simulations to be scaled, enabling large or complex simulations, that otherwise wouldn't have been able, to run in real-time.
	\item Containerisation of Game Engines. Game engines, generally, currently run as one single application, including graphics, physics, AI etc. (with the exception of databases for user data). AP separates out the physics element into its own container, much like many modern systems are splitting up different aspects of systems into their own containers, for example web applications. This containerisation of physics, means, physics can be elastically scaled independently from other systems such as graphics, along with other advantages of modularity.
	\item Streamed Gaming. The main foreseeable application of AP is Streamed Gaming, with the use of large multiplayer worlds, that are persistent and scalable with dedicated and elastic physics nodes. Physics is no longer limited to what can be simulated on a single machine and co-located render nodes solves the issue of higher latencies between the render node and physics simulation, the main problem with the use of physics in online gaming.
\end{itemize}

>>>>

\chapter{Introduction}
%TODO: Introduction paragraph?
\section{Real-Time Physics}

Physics has been a part of video games for almost as long as video games have existed \cite{millington2007game}. Real-time physics is necessary to provide realistic 2D and 3D environments to players as it allows players to interact with the simulated environment. For example, moving crates around, simulating the flight of a plane or the ballistics of a bullet. Physics engines are responsible for all of the calculations required to simulate physics. Physics engines for games aim to achieve `realism', but sacrifice accuracy in order to achieve real-time speeds. Recently, the mathematical error from commercial physics engines has been reduced and physics engines are finding use in other simulations besides video games (e.g.,  \cite{Xu2017, Lu2017, Shah2018}).

%\section{Distributed Real-Time Physics}
%Physics is important for games, it is used for...
%Video games rely on real-time physics engines to provide realistic environments. This is noticeable in 3D gaming where players expect fast paced interaction. This requirement has produced significant commercial activity resulting in physics engines for real-time simulation for rigid body dynamics. Although realism is convincing, a degree of mathematical accuracy is usually sacrificed (causing error) in order to attain real-time delivery within constrained resource settings. However, with significant commercial effort, the mathematical error has reduced over the years and commercial game engines are finding their use in other industries (e.g.,  \cite{Xu2017, Lu2017, Shah2018}).

%Combining physics engines with game engines (e.g., Unity, Unreal), presents a cost-efficient development platform for developing realistic environments. These gaming environments may be presented to geographically distant players using server-side scalable technologies. This allows scalable gaming platforms, presenting players with large online worlds to explore whilst maintaining the visually highly realistic environments generated by the player's local machine.

\section{Scalable Games and Simulations}
Scaling and more complex games is a selling point. This has been demonstrated by games such as EVE Aether Wars \cite{AetherWars} and \cite{WorldsAd48:online}, built using HadeanOS and SpatialOS respectively. HadeanOS and SpatialOS both provide attempts at distributing physics by allowing physical interactions across servers. Despite the growing commercial interest in this area, there is little academic literature into the challenges and solutions of distributed real-time physics.

Research into delivering scalability in such online worlds is focused on balancing real-time and consistency issues for enabling player-player interaction. The evolution of such research can be traced back to the earliest work on game area subdivision \cite{Npsnet} through player-focused game area sub-division (e.g., \cite{Morgan2005InterestManagement, Greenhalgh}) and eventually to the many commercial cloud based solutions as described by \cite{iCloudAccess}.

Traditionally real-time simulations are confined to a single machine, including real-time physics. 
Recent cloud technology has made available, lower cost, low-latency, high-performance, including GPU-enabled machines. This means, real-time simulations can now run in the cloud as well as being economically viable. It is possible to rapidly provision resources on demand, allowing for fast scaling up and down of resources as needed.

Physics is the most challenging problem facing distributing games engines. Handling player-player interaction between servers, which is still not a solved problem, requires response times in the order of hundreds of milliseconds and still leads to inconsistencies in the game world. Real-time physics solvers work in iterations normally measured in the order of less than 10 milliseconds. Although this poses a significant challenge, achieving distributed real-time physics would allow the physics engine component of the game engine to be elastically scaled, meaning the physics engine is no longer limited in scale and complexity to what would be achievable on a single machine.

%The problem of scalable player-player interaction is not a solved problem, as the numerous interaction possibilities across gaming genres and the latency in networks provides many different possible bespoke solutions. Player interaction can be measured in milliseconds, which current networking technology cannot model for all gaming genres in a scalable manner without giving rise to inconsistency. Therefore, the notion of trying to model real-time physics, which requires solvers working in iterations measured in microseconds, would be considered a near impossible task. However, achieving this would allow the physics engine element of the gaming console to be transplanted to the server-side (cloud), freeing up resources for game-play and rendering on a player's local machine.

@@

\section{Streamed Gaming}
Streamed gaming is an active area of research (e.g., \cite{GARCIAVALLS}), where the game is executed on the server-side, the graphical output is streamed to a remote player and input is sent from the player to the server-side.
Streamed gaming reduces the computing power requirements on player hardware, even allowing gaming on mobile devices with graphical quality beyond what would be possible if the game were running entirely on the device \cite{CloudGamingArchPerf}. In addition, this removes the need for players to continuously upgrade their hardware \cite{LatencyoOfCloudGamingSystems}.

The most difficult challenge in streamed gaming is the quality of experience of games, as the experience is significantly impacted by delay\cite{StormInCloudGaming}. The effect on quality depends upon the time of game, for example low-latency games like first person shooters require response times as little as 100ms, whereas, real-time strategy games can have acceptable response times of up to 1000ms \cite{CloudGamingArchPerf}. Recent advances in cloud gaming, including the use of edge cloud game servers, have enabled the cloud to be utilised to deliver low-latency gaming \cite{TimelyCloudGaming}.

Examples of streamed gaming services include Sony's PSNOW \cite{PSNOW}, NVidia's GameStream \cite{NVidiaGameStream} and Google's Stadia \cite{Stadia}. The existence of such services suggests the possibility of real-time interaction for all gaming genres across cloud infrastructures.

%Recent developments have demonstrated the importance of software defined networks (SDN) for presenting timely cloud solutions \cite{Wang, Kumar}. The SDN manages the distribution of messages across servers in the most economical way possible (e.g., resource usage). The fact that SDNs can produce timely messaging across tightly coupled server-side deployments (the cloud) suggests the possibility that timely requirements for scalable physics delivery can be fulfilled.

%If the physics simulation is co-located with the graphics streaming servers (with very low-latency), the problem of latency between player's view of the physics simulation and actual state is resolved, allowing for a better shared view of the physics state.

\section{Containerisation}
It therefore is the next logical step to start breaking apart the different systems of the game into their own 'containers', allowing for independent scaling and resource provision management. This would allow for a more cost efficient approach to scalable and streamed games and simulations. Our problem is concentrated on distributing the physics engine element of game engines across the server-side while maintaining the real-time streamed service to players.
Using the recent developments in cloud technology, we explore the possibility of using this on-demand scaling, to scale up real-time physics simulations.

\section{Summary}
%TODO: Rewrite this section
In the context of current research, we can place ourselves at the juncture of balancing physics calculations across servers in the cloud for game streaming services. This should provide improved economic use of resources, as current game streaming services simply mimic the machine requirements of the player without distributing any of the component parts of the gaming engine. In addition, streamed games may benefit from increased numbers of artefacts (way beyond what is currently possible in a player's machine) as other servers may be utilised to solve the physics problems.

@@

\section{Outline}
%TODO: Make sure this is up to date
The following section provides an outline for this study:
2. Background and Related Work:
\begin{itemize}
	%Context
	\item Real-time physics engines overview. We introduce the reader to the topic of physics simulations and we compare real-time with non-real-time physics simulations. We define terms, workings, important and relevant aspects, e.g. body types, uses in games and simulations, typical scales and performance of real-time simulations and performance budget in real-time applications.
	\item Scalable Non-Real Time Physics. We discuss previous work and applications of scalable non-real time physics and how those techniques are not suitable for use in real-time physics simulations.
	\item Online gaming. We discuss the wider topic of online gaming (client-hosted, P2P, server), challenges of latency, security/cheating and computational limits. Also some details of online gaming servers are discussed, specifically, typical server frame-times and network tick updates. 
	\item Physics in online gaming (and the lack of, due to the challenges faced).
	%Relevant Related Work
	\item Graph-partitioning. When it comes to spatially-partitioning a simulation, the problem is essentially a graph partitioning problem, where objects are nodes and if there’s a possible interaction, an edge exists. Useful for terms, comparisons with and future work on load-balancing will heavily involve this.
	\item Distributed virtual environments. We look at existing ways of distributing game worlds and previous work carried out on these approaches. These fall into two categories, migratory (approach used by AP) and non-migratory. The studies found non-migratory to be the better solution when focusing on player-player interaction and we will discuss why migratory is the better solution for physics body-physics body interaction. Important concepts such as interest management are introduced to the reader.
	\item Distributed Real-time Physics. We will discuss previous work looking specifically at distributing real-time physics and how AP differs from these approaches.
	\item Cloud Computing. We discuss the wider context of cloud computing, how AP is deployed on the cloud and also specifically conditions that affect AP in the cloud, i.e. latency and packet-loss
	%Applications
	\item Scientific and Engineering simulations. Real-time simulations are being used more in scientific research, however these are limited to what can be achieved on a single machine. AP allows real-time physics simulations to be scaled, enabling large or complex simulations, that otherwise wouldn't have been able, to run in real-time.
	\item Containerisation of Game Engines. Game engines, generally, currently run as one single application, including graphics, physics, AI etc. (with the exception of databases for user data). AP separates out the physics element into its own container, much like many modern systems are splitting up different aspects of systems into their own containers, for example web applications. This containerisation of physics, means, physics can be elastically scaled independently from other systems such as graphics, along with other advantages of modularity.
	\item Streamed Gaming. The main foreseeable application of AP is Streamed Gaming, with the use of large multiplayer worlds, that are persistent and scalable with dedicated and elastic physics nodes. Physics is no longer limited to what can be simulated on a single machine and co-located render nodes solves the issue of higher latencies between the render node and physics simulation, the main problem with the use of physics in online gaming.
\end{itemize}
